---
title: "Report 7 2.0: How does media coverage change through time?"
subtitle: "A casual report"
author: "Ricky Heinrich for Mélanie Méthot"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: true
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F, message = F, warning = F,fig.pos = "H", fig.height = 4.3)
```
```{r}
library(readr)
library(anytime)
library(stringr)
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr)
library(reshape2)
library(tidyr)
library(gridExtra)
library(kableExtra)
library(zoo) # for rollmean
```

```{r}
# had to go back into wrangling_cleaning and add # of articles column
states <- read_csv("../data/states.csv", show_col_types = FALSE)
# setting colors for sentences
colors_sentences <- c("lightgrey","darkgrey","#FBE6A2","#EECDCD","#DFBAB1", "#D08370", "#BD4B31","#982b15", "#7a2917","#531607")
```


<!-- first, I can look at a simple table to see what's going on. `r kable(table(states$num_articles))` looks like not all just plain numbers - some have words. Could start with converting to numerical, assuming that any that don't work will end up as NA, then looking into those that are NA and seeing if they need fixes, or send to mm for review.-->

# what's new
- (in bold in text)
- `r round(nrow(filter(states, is.na(states$num_articles)))/nrow(states)*100,1)`% of the cases in the corpus don't have anything recorded for "# of articles".
- found in some cases, says '0' in "# of articles" but then there is something in link1 ...
- changed categories after sept 2 meeting: remove 0 as large proportion due to 'not done yet', 1, 5-9, 10-29 (public interest), 30+ (sensation); check 'categories' subtitle
- check coverage through time -> decades
- coverage by states w/o incomplete -> check analysis


# Overview

It looks like there are `r nrow(filter(states, is.na(states$num_articles)))` cases that don't have the number of articles recorded. **We are assuming that these cases are not yet completed yet, ie areas where research still needs to be done. These are `r round(nrow(filter(states, is.na(states$num_articles)))/nrow(states)*100,1)`% of the cases in the corpus. These will not be included in the analysis.**

The cases with empty cells for number of articles occur across five of the states, shown in the table below. A quick review in the original spreadsheet of Queensland, Western Australia, and NSW  shows that at least some of these cases do have links. The lack of value in this cell thus comes from a deficiency in the transcriber. The two cases in South Australia don't have any links, and the first few in Victoria also don't, but I didn't look at all of the cases. 

```{r}
kable(table(filter(states, is.na(states$num_articles))$state))
```

I don't have a quick an easy way right now to check the number of links and input a value, although I am sure it is doable. I may try to think about it and find a way later. 

~~There are `r nrow(filter(states, is.na(states$link1) & is.na(states$num_articles))[,c("state","link1")])` cases where no entry was recorded for "# of articles" AND there was nothing recorded in "link1". I am adding the entry '0' in those cases.~~ 

**There are `r nrow(filter(states, !is.na(states$link1) & states$num_articles == '0'))` cases where '0' was recorded for "# of articles" AND there was something recorded in "link1". Annoying. I'm putting it under insufficient research for now.**

<!-- Could be more straightforward to at least check if there is something only in the Link1 cell; if empty, can assume no links, if not, at least 1 link. Will think about it. Then at least I could tally up how many are have no links recorded so no coverage, vs how many are just transcriber deficiencies. Maybe I will try to do think - add Link1 column, then if num_articles and link1 are NA, then make num_articles == "no link1", if not, then "mistake" or "at least 1"-->




<!-- If there's yorick in the cell, then count as 'yorick?' or just delete-->
```{r}
#filter(states, !is.na(states$link1) & is.na(states$num_articles))[,c("state","link1")]
# convert to NA if has link but says 0
states$num_articles <- ifelse(!is.na(states$link1) & states$num_articles == '0', NA, states$num_articles) 

# convert to numeric; if not just a number then will be NA
states$articles_numeric <- as.numeric(states$num_articles)


# if 'yorick' in link1, then delete and count as empty cell -- I checked in og spreadsheet and there were never links in link2
states$link1 <- tolower(states$link1) # convert all to lower; idc about actually accessing link so its ok
#states$link1 <- sub(".*yorick.*", NA, states$link1)
```


<!-- IMPORTANT!! "HARD CODED" A VALUE WHEN I SHOULDN:T HAVE BUT WAS LAZY (100 - xxx) in previous r code in text. -->
There are `r nrow(filter(states, !is.na(states$link1) & is.na(states$num_articles) &!grepl('yorick', states$link1)))` cases where there's no entry in the number of articles column, but there is one link or what seems like an article title. `r nrow(filter(states, !is.na(states$link1) & is.na(states$num_articles) & grepl('yorick', states$link1)))` cases had something along the lines of "yorick's info" but didn't include any links, so for now have remained as having no entry for # of articles. Since there there isn't an easy straightforward way to count how many links there are per cases, and a quick overview showed that some included more than 5 links, I will not change these values to '1' article as it would skew the results. I will instead just not include them in the analysis, and leave them as 'no entry', which is valid since it is a small proportion of the data. Maybe later I'll go back and take the time to manually add the number of links since I guess it would be manageable.

There are a few cases where the entry in "# of articles" is not just a numerical digit, shown in the table below. For those that have a variation of 'more than x', I'll just put in the x. For 'cannot find articles' and 'none' I will put '0'.
<!-- next look at cases that weren't NA in num_articles but are in articles_numeric -->
```{r}
kable(table(filter(states, is.na(states$articles_numeric))$num_articles))
```


```{r}
states$num_articles <- sub("none", 0, states$num_articles)
states$num_articles <- sub(".*cannot find.*", 0, states$num_articles)
states$num_articles <- stringi::stri_extract_first_regex(states$num_articles, "[0-9]+")

# need to rerun this
# convert to numeric; if not just a number then will be NA
states$articles_numeric <- as.numeric(states$num_articles)

# convert if no links and no entry -> 0
#states$articles_numeric <- ifelse(is.na(states$link1) & is.na(states$num_articles), 0, states$articles_numeric)

states$link1 <- sub(".*yorick.*", NA, states$link1)
```

# Count Bar Plot
The maximum number of articles is `r max(states$articles_numeric, na.rm = T)`. The plot below shows the count of article numbers, and the two after show it zoomed in. **We see that the most occurring is '3', which occurs `r nrow(filter(states, states$articles_numeric == 3))` times. ** There's somewhat of a peak at 3 articles, then a somewhat gradual decline, with small peaks at 11, 16, 17, 22, 25, ... these peaks are relatively small differences: the number of cases with 10 articles is 82, and that with 11 is 87.



```{r}
# histogram
ggplot(states) + geom_histogram(aes(x = articles_numeric), color="black", fill="white", binwidth = 1)+ labs(title = "Count of Media Coverage per Case", x="number of articles") 

ggplot(states) + geom_histogram(aes(x = articles_numeric), color="black", fill="white", binwidth = 1) + xlim(-1,100)+ labs(title = "Count of Media Coverage per Case", x="number of articles")

ggplot(states) + geom_histogram(aes(x = articles_numeric), color="black", fill="white", binwidth = 1) + xlim(-1,50) + ylim(-1, 230) + labs(title = "Count of Media Coverage per Case", x="number of articles")

#ggplot(states) + geom_histogram(aes(x = articles_numeric), color="black", fill="white", binwidth = 1) + xlim(27,101) + ylim(-1, 50) + labs(title = "Count of Media Coverage per Case", x="number of articles")
```
Here is the exact numbers in case you wanted to refer to them directly; every second row is the count. 

```{r}
table(states$articles_numeric)
```

# Categories

**After our discussion on Sept 2, we are categorizing as: 0, 1, 2-4, 5-9, 10-29, 30+. The following plot shows how many cases fall in each category. The most occuring is 10-29 (public interest).**

```{r}
# 0, 1-3, 4-9, 10-15, 16-27, 28-69, and 70+. 
# new: 0, 1, 2-4, 5-9, 10-29, 30+
# change NA to -1
states$articles_numeric <- replace_na(states$articles_numeric, -1)
group_boundaries_numart <- c(-1,0, 1, 2,5,10,30, max(states$articles_numeric))
# -1,0, 1, 4,10,16,28,70
states$group_numart  <- findInterval(states$articles_numeric, group_boundaries_numart, all.inside = T)

states$group_numart <- factor(states$group_numart, labels = paste0(str_sub(group_boundaries_numart, 1, 4)[-length(group_boundaries_numart)],"-",as.numeric(str_sub(group_boundaries_numart, 1, 4)[-1])-1))

ggplot(data.frame(table(states$group_numart)), aes(x=Var1, y = Freq)) +
  geom_col()+ geom_text(aes( label = Freq),position = position_stack(vjust = .5), color = "white")+ labs(title="Count of Cases per Media Coverage Category", x = "# of Articles", y = "Count") + 
          theme(
            axis.text.x = element_text(size=10, face="bold", color = "#333333"),
            axis.text.y = element_text(size=10, face="bold", color = "#333333"),
            legend.text = element_text(size=10, face="bold", color = "#333333"),
            legend.title = element_text(size=13, face="bold", color = "#333333"),
            title = element_text(size=12, face="bold", color = "#333333"))
  
  

```


# Coverage Through Time
Next we have these groups in time.**We see that the groups with the largest difference in media coverage through time is the 1-4: there is an increase of cases with only 1-4 articles near the WWII years. The following plot is by decade: we see that there's a pretty drastic increase in cases with 2-4 articles in the 1940s, with 200 cases. There's also increases across the decades in cases with 5-9 articles and 10-29, although both have peaks that occured somewhat earlier: case with 5-9 articles peaked at around 110 in the 1930s, whereas 10-29 peaked jointly in 1910s and 1920s at around 125. There were only about 50 cases in the 1940s with 10-29 articles. Sensations (30+ articles) occured very infrequently in those years: sensations appeared to have peaked in the 1890s, with nearly 50 cases.** 

```{r fig.height=10, fig.width=7}
ggplot(states) + geom_line(aes(year(Date), colour=factor(group_numart)), stat = "count") + labs(x="", title = "Count by 'number of articles' groups through time",subtitle = "-1--1 represents unfinished cases", colour="") +facet_wrap(~group_numart, ncol = 2, scales = "free") + scale_x_continuous(n.breaks = 16, limits = c(1810,1970))+ scale_y_continuous(limits = c(0,75))+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), legend.position = "none")

#ggplot(states) + geom_point(aes(year(Date), colour=factor(group_numart)), size = 1, stat = "count") + labs(x="", title = "Count by 'number of articles' groups through time",subtitle = "-1--1 represents unfinished cases", colour="") +facet_wrap(~group_numart, ncol = 2)+ scale_x_continuous(n.breaks = 16, limits = c(1810,1970))+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# by decade
states %>% mutate(decade = as.numeric(as.character(year(Date)))-as.numeric(as.character(year(Date)))%% 10) %>% ggplot()+geom_bar(aes(x = decade, fill = group_numart))  + 
  labs(x="", title = "Count by 'number of articles' groups by Decade",subtitle = "-1--1 represents unfinished cases", colour="", )+
  facet_wrap(~group_numart, ncol = 2, scales = "free")+
  scale_x_continuous(n.breaks = 16, limits = c(1810,1970))+
  scale_y_continuous(limits = c(0,500))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), legend.title = element_blank(), legend.position = "none")

```

```{r eval=F}
# rug plot
ggplot(states,aes(x = Date,y = group_numart )) +  
  geom_point(shape="|",size=4) + labs(title = "Distribution of judges' sentence proportions", x = "Proportion", y = "Sentence") 
```

# Coverage by State

## With Incomplete

**The first set of plots here include incomplete cases. I looks like a third of NSW and Queensland cases are incomplete (-1--1).  **

```{r}
ggplot(states) + geom_bar(aes(y = state, fill=factor(group_numart))) + labs(x="Count", title = "Count by 'number of articles' groups by states",subtitle = "-1--1 represents unfinished cases", colour="", y="")+ scale_fill_manual(values = colors_sentences)+theme(legend.title=element_blank())

ggplot(states) + geom_bar(aes(y = state, fill=factor(group_numart)), position = "fill") + 
  labs(x="", title = "Proportion by 'number of articles' groups by states",subtitle = "-1--1 represents unfinished cases", colour="", y= "")+
  scale_x_continuous(labels = scales::percent)+ 
  scale_fill_manual(values = colors_sentences)+theme(legend.title=element_blank())
```

## Without Incomplete

**There are more cases in Victoria with 2-4 articles than in NSW, and it looks like Victoria has somewhat more cases than NSW that were sensations (30+ articles). Queensland seems to have an equal amount of cases with 5-9 articles than with 10-29 articles. Queensland also has a very small number of cases with only 1 article (3), even less than South Australia (5) (Exact number found in table below). You can see that the sensations (30+ articles) occurs in every state. **

```{r}
ggplot(filter(states, articles_numeric != -1)) + geom_bar(aes(y = state, fill=factor(group_numart))) + labs(x="Count", title = "Count by 'number of articles' groups by states", colour="", y= "")+ scale_fill_manual(values = colors_sentences[-1])+theme(legend.title=element_blank())

ggplot(filter(states, articles_numeric != -1)) + 
  geom_bar(aes(y = state, fill=factor(group_numart)), position = "fill") + 
  labs(x="", title = "Proportion by 'number of articles' groups by states", colour="", y="")+
  scale_x_continuous(labels = scales::percent)+ 
  scale_fill_manual(values = colors_sentences[-1])+
  theme(legend.title=element_blank())
```

The exact numbers are shown in the table below. 
```{r}
kable(table(states$state, states$group_numart))
```

The following shows the proportions for easier quantifying. For example, the value intersecting 2-4 and NSW 12.3 means that 12.3% of NSW's cases have article counts between 2 and 4. 

```{r}
kable(round(prop.table(table(states$state, states$group_numart), margin = 1)*100, 1))
```

