---
title: "Report 5 4.1:Judges & Sentences by states"
subtitle: "A casual report"
author: "Ricky Heinrich for Mélanie Méthot"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: true
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F, message = F, warning = F,fig.pos = "H")
```
```{r}
library(readr)
library(anytime)
library(stringr)
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr)
library(reshape2)
library(tidyr)
library(gridExtra)
library(kableExtra)
```

```{r}
states <- read_csv("../data/states_clean.csv", show_col_types = FALSE)

# setting colors for sentences
colors_sentences <- c("lightgrey","darkgrey","#FBE6A2","#EECDCD","#DFBAB1", "#D08370", "#BD4B31","#982b15", "#7a2917","#531607")

# capitalizing judge names
states <- states %>% mutate(judge_simple = str_to_title(judge_simple))

```

# What's new
- change from proportions to %
- insight: Victoria % of 'yes prison' never go below 50%; all of Victoria top 15 judges sentence at least 50% to prison. other states at least some of the judges sentence smaller proportions. 
- edited requested plots to be more presentation friendly
- add judge timelines (found first plot after state header)
- add war years pie charts: before and during WWI have very similar proportions, tend towards more people 'yes prison'. between the wars there's a change, slightly more 'no prison' than 'yes prison'. during WWII similar. after WWII more 'no prison' than 'yes prison'. 

# War Years

```{r}

# add convict (yes/no) column
states <- states %>%  mutate(convict = ifelse(sentence_simple == "other" | sentence_simple == "D.N.S." | sentence_simple == "unknown", "no","yes" ))

# count bar chart, highlight war years
ggplot(states, aes(x = year(Date)))  +
  geom_rect(aes(xmin = 1914, xmax = 1919,ymin = -Inf, ymax = Inf), alpha = 0.0099, fill = "#ff6666",show.legend = F) +
  geom_rect(aes(xmin = 1939, xmax = 1946,ymin = -Inf, ymax = Inf), alpha = 0.0099, fill = "#ff6666",show.legend = F)+ geom_bar() + labs(title = "Count of Cases Per Year, War Times Highlighted", x = "", subtitle = "WWI: 1914-1919, WWII: 1939-1946")+ scale_x_continuous(n.breaks = 16)+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# before WWI
ggplot(filter(states,  (year(Date) < 1914)), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "Before WWI (-1914)", fill = "Prison?", subtitle = paste("n =",nrow(filter(states,  (year(Date) < 1914))))) +
  scale_fill_manual(values = colors_sentences[3:4])

# pie chart WWI
ggplot(filter(states, year(Date) >= 1914 & year(Date) <= 1919), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "During WWI (1914-1919)", fill = "Prison?", subtitle = paste("n =", nrow(filter(states, year(Date) >= 1914 & year(Date) <= 1919)))) +scale_fill_manual(values = colors_sentences[3:4])

# pie chart b/w the wars
ggplot(filter(states,  (year(Date) > 1919 & year(Date) < 1939)), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "Between the wars (1920-1938)", fill = "Prison?", subtitle = paste("n =", nrow(filter(states,  (year(Date) > 1919 & year(Date) < 1939))))) +scale_fill_manual(values = colors_sentences[3:4])

# pie chart WWII
ggplot(filter(states, year(Date) >= 1939 & year(Date) <= 1946), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "During WWII (1939-1946)", fill = "Prison?", subtitle = paste("n = ", nrow(filter(states, year(Date) >= 1939 & year(Date) <= 1946)))) +scale_fill_manual(values = colors_sentences[3:4])

# pie chart after WWII
ggplot(filter(states, year(Date) > 1946), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "After WWII (1947-)", fill = "Prison?", subtitle = paste("n = ", nrow(filter(states, year(Date) > 1946)))) +scale_fill_manual(values = colors_sentences[3:4])

# pie chart All not war times
ggplot(filter(states, year(Date) < 1914 | (year(Date) > 1919 & year(Date) < 1939) | year(Date) > 1946), aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = "All not war times (-1913, 1920-1938, 1947-) ", fill = "Prison?", subtitle = paste("n =", nrow(filter(states, year(Date) < 1914 | (year(Date) > 1919 & year(Date) < 1939) | year(Date) > 1946)))) +scale_fill_manual(values = colors_sentences[3:4])


# pie chart overall
ggplot(states, aes(x = "", fill = convict))  +
  geom_bar(position = "fill") + geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_fill(vjust = 0.5)) +
  coord_polar("y", start = 0, direction = -1)+ theme_void()+
  labs(title = paste0("All time (", min(year(states$Date)), "-",max(year(states$Date)),")") , fill = "Prison?") +scale_fill_manual(values = colors_sentences[3:4])


```

```{r}
list_state <- unique(states$state)
num_state = 1

judge_counts_prop <- function(num_state){
  # filter for current state
  fcn_df_all <- filter(states, state == list_state[num_state]) 
  
  # add convict (yes/no) column
  fcn_df_all <- fcn_df_all %>%  mutate(convict = ifelse(sentence_simple == "other" | sentence_simple == "D.N.S." | sentence_simple == "unknown", "no","yes" ))
  # filter only top 15 judges
  table_judge <- table(fcn_df_all$judge_simple)
  fcn_df <- filter(fcn_df_all, judge_simple %in%  names(sort(table_judge, decreasing = T))[1:15])
  # getting judge year ranges: group by judge, get min and max years
  judge_years <- fcn_df %>% group_by(judge_simple) %>% 
    summarise(first_year = min(year(Date)), last_year = max(year(Date)), 
              range = last_year - first_year)
  # create new col with merged info
  
  judge_years$judge_info <- paste0(judge_years$judge_simple," (",
                                   judge_years$first_year,"-",
                                   judge_years$last_year, ")")
  # judge_years$judge_info <- paste0(judge_years$judge_simple," ",
  #                                  judge_years$first_year,"-",
  #                                  judge_years$last_year, " (",
  #                                  judge_years$range," yrs)")
  
    # judge_years$judge_info <- paste0(judge_years$judge_simple," (",
    #                                judge_years$range," yrs)")
  
  # merge on judge_simple, add judge_info only
  fcn_df <- merge(x = fcn_df, 
                  y = judge_years[ , c("judge_simple", "judge_info")],
        by = "judge_simple", all.x=TRUE)
  
  
  # table frequency, clean new df
  fcn_freq <- data.frame(table(fcn_df$judge_info, fcn_df$convict))
  fcn_freq <- dcast(fcn_freq, Var1~Var2, value.var = "Freq")
  colnames(fcn_freq)[colnames(fcn_freq) == "Var1"] <- "judge"
  # adding sum, proportions
  fcn_freq$sum <- rowSums(fcn_freq[,-1])
  fcn_freq <- fcn_freq %>% mutate(across(2:3, ~./sum, .names = "{col}_prop"))
  
  # sort by decreasing total
  #fcn_freq <- fcn_freq[order(-fcn_freq$sum),]
  # sort by decreasing proportion
  fcn_freq <- fcn_freq[order(-fcn_freq$yes_prop),]
  row.names(fcn_freq) <- NULL
  # order plot by frequency
  fcn_df$judge_info <- factor(fcn_df$judge_info, levels = fcn_freq$judge)
  
  #fcn_melted <- melt(judge_years[,c(2,3,5)])
  #fcn_melted$judge_info <- factor(fcn_melted$judge_info, levels = fcn_freq$judge)
  
  # + geom_point(aes(x=value, y = judge_info))
  # print(ggplot(fcn_melted) +
  #         geom_line(aes(x=value, y = judge_info, color = judge_info)) +
  #         theme(legend.position = "none"))
  
  # judges timeline plot
  print(ggplot(fcn_df) +
          geom_line(aes(x = year(Date), y = judge_info, color = judge_info), size = 2) +
          scale_x_continuous(n.breaks = 16) +
          labs(title=paste(list_state[num_state], "Judges Time Ranges"), x = "", y = "") +
          theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), legend.position = "none")+ 
          theme(
            axis.text.x = element_text(size=10, face="bold", color = "#333333"),
            axis.text.y = element_text(size=10, face="bold", color = "#333333"),
            legend.text = element_text(size=10, face="bold", color = "#333333"),
            legend.title = element_text(size=13, face="bold", color = "#333333"),
            title = element_text(size=12, face="bold", color = "#333333")))
  ggsave(paste0("report5_4.2_judgetime",list_state[num_state], ".png"))
  # +geom_point(aes(x = year(Date), y = judge_info), shape="|",size=2)  # good for cleaning/exploring purposes, not for presentation b/c raises questions of having done it right or not ... for now can remove, open to 'accusations' of mistakes
  
  # proportion plot
  print(ggplot(fcn_df, aes(y=judge_info, fill = convict)) + 
          geom_bar(position = "fill") +
          annotate("text", x = -0.02, y = seq_along(fcn_freq$sum), 
                   label = fcn_freq$sum,vjust = 0.5, hjust = +0.7, size = 3) + 
          scale_fill_manual(values = colors_sentences[3:4]) + 
          labs(title = paste(list_state[num_state], "Sentencing Proportions"), x = "", y = "", fill = "Prison?", subtitle = "Top 15 Judges")+scale_x_continuous(labels = scales::percent)+ 
          theme(
            axis.text.x = element_text(size=10, face="bold", color = "#333333"),
            axis.text.y = element_text(size=10, face="bold", color = "#333333"),
            legend.text = element_text(size=10, face="bold", color = "#333333"),
            legend.title = element_text(size=13, face="bold", color = "#333333"),
            title = element_text(size=12, face="bold", color = "#333333")))
  ggsave(paste0("report5_4.2_",list_state[num_state], ".png"))
  
  # +  geom_vline(xintercept = mean(fcn_freq$yes_prop),  linetype = "dashed", color = "red") 
  
  cat(paste0("\nThe proportion range for ",list_state[num_state], " is ", round(min(fcn_freq$yes_prop)*100,1), "% to ", round(max(fcn_freq$yes_prop)*100,1), "%.\n"))
  
  # count plot
  print(ggplot(fcn_df, aes(y=judge_info, fill = convict)) + geom_bar()+
    annotate("text", x = -0.02, y = seq_along(fcn_freq$yes_prop), label = paste0(round(fcn_freq$yes_prop*100,1),"%"),vjust = 0.5, hjust = -0.1, size = 3)+ expand_limits(x = c(0,50))+
    scale_fill_manual(values = colors_sentences[3:4]) +
    labs(title = paste(list_state[num_state], "Sentencing Counts of Top 15 Judges"), x = "Count", y = "", fill = "Prison?")) 
  
  # in time
  print(ggplot(fcn_df_all) + geom_bar(aes(x=as.numeric(as.character(year(Date))), fill = convict), position ="identity", alpha = 0.7) + labs(title = paste("Sentencing Counts Through the Years in",list_state[num_state] ), x = "", subtitle = "All judges", fill = "Prison?")+ scale_x_continuous(n.breaks = 16)+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)))
}

# want to add average and std dev of proportions, and maybe counts?
#mean(df_freq$D.N.S._prop)
#sd(df_freq$D.N.S._prop)

#print(paste("The proportion range for",list_state[num_state], "is", min(df_freq$D.N.S._prop), "to", max(df_freq$D.N.S._prop), "."))


# if want a timeline: want a line connecting start of residency anad at end of residency, x = years, y = judge


#ggplot(melt(judge_years[1:15,1:3])) + geom_point(aes(x=value, y = judge_simple))+ geom_line(aes(x=value, y = judge_simple, color = judge_simple))

#ggplot(states[15:30,]) + geom_line(aes(x=year(Date), y = judge_simple, color = judge_simple))

```

```{r}
num_state = 1
```
# `r list_state[num_state]`

The next 3 plots are repeated for each state. The first plot shows the proportion of prison vs not imprisoned for the top 15 judges for that state, and the number beside each name is the count of cases for that judge. The second plot shows the specific counts, and the number beside each name is the proportion of cases where the accused was sentenced to prison time. The third plot shows the counts per year for all judges.

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r}
judge_counts_prop(num_state)
num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges.

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```




