---
title: "Report 5 4.1:Judges & Sentences by states"
subtitle: "A casual report"
author: "Ricky Heinrich for Mélanie Méthot"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    toc: true
    extra_dependencies: ["float"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, cache = F, message = F, warning = F,fig.pos = "H")
```
```{r}
library(readr)
library(anytime)
library(stringr)
library(ggplot2)
library(lubridate)
library(dplyr)
library(knitr)
library(reshape2)
library(tidyr)
library(gridExtra)
library(kableExtra)
```

```{r}
states <- read_csv("../data/states_clean.csv", show_col_types = FALSE)

# setting colors for sentences
colors_sentences <- c("lightgrey","darkgrey","#FBE6A2","#EECDCD","#DFBAB1", "#D08370", "#BD4B31","#982b15", "#7a2917","#531607")

# capitalizing judge names
states <- states %>% mutate(judge_simple = str_to_title(judge_simple))

```

# What's new
- capitalize judge names
- moved proportions number further right on count plots
- added active years info by judge names
- sort by the proportions of 'yes' prison
- added a red line showing the mean proportion for the top 15 judges of that state (will remove if you don't find value in it)
- scale some states count plot to same x-axis (0-50) for easier comparison (all except vic and NSW that go over it). depending on your presentation we can change this; want them somewhat constant if doing direct comparisons, so that people aren't misled into thinking states with lower counts have more than they actually do.



# Notes
I want to reformat so that each same plot for all sides is printed alongside each other to ease comparison. Doesn't currently work with the way I formatted the code, so I need to come up with something different to do so. Wanted to provide you with this in the meantime, just have to scroll to compare between states. I haven't yet taken the time to analyse each, we can go over it together if you'd like so you can ask questions simultaneously.

Let me know of changes you'd like (can choose different colors, change titles for example), and which you'd like sent as pngs for use. I could also change the order in which the judges are presented. Currently they are in order of count of cases each have, but they could be in order of prison proportions or prison counts (the ones with largest amount of 'yes' prison come first).

My to do list:

- look into number of articles per state, per judge, per time period
- Answer: does newspaper coverage change over time? overall, per state

```{r}
list_state <- unique(states$state)
num_state = 1

judge_counts_prop <- function(num_state){
  # filter for current state
  fcn_df_all <- filter(states, state == list_state[num_state]) 
  
  # add convict (yes/no) column
  fcn_df_all <- fcn_df_all %>%  mutate(convict = ifelse(sentence_simple == "other" | sentence_simple == "D.N.S." | sentence_simple == "unknown", "no","yes" ))
  # filter only top 15 judges
  table_judge <- table(fcn_df_all$judge_simple)
  fcn_df <- filter(fcn_df_all, judge_simple %in%  names(sort(table_judge, decreasing = T))[1:15])
  # getting judge year ranges: group by judge, get min and max years
  judge_years <- fcn_df %>% group_by(judge_simple) %>% 
    summarise(first_year = min(year(Date)), last_year = max(year(Date)), 
              range = last_year - first_year)
  # create new col with merged info
  
  judge_years$judge_info <- paste0(judge_years$judge_simple," (",
                                   judge_years$first_year,"-",
                                   judge_years$last_year, ")")
  # judge_years$judge_info <- paste0(judge_years$judge_simple," ",
  #                                  judge_years$first_year,"-",
  #                                  judge_years$last_year, " (",
  #                                  judge_years$range," yrs)")
  
    # judge_years$judge_info <- paste0(judge_years$judge_simple," (",
    #                                judge_years$range," yrs)")
  
  # merge on judge_simple, add judge_info only
  fcn_df <- merge(x = fcn_df, 
                  y = judge_years[ , c("judge_simple", "judge_info")],
        by = "judge_simple", all.x=TRUE)
  
  
  # table frequency, clean new df
  fcn_freq <- data.frame(table(fcn_df$judge_info, fcn_df$convict))
  fcn_freq <- dcast(fcn_freq, Var1~Var2, value.var = "Freq")
  colnames(fcn_freq)[colnames(fcn_freq) == "Var1"] <- "judge"
  # adding sum, proportions
  fcn_freq$sum <- rowSums(fcn_freq[,-1])
  fcn_freq <- fcn_freq %>% mutate(across(2:3, ~./sum, .names = "{col}_prop"))
  
  # sort by decreasing total
  #fcn_freq <- fcn_freq[order(-fcn_freq$sum),]
  # sort by decreasing proportion
  fcn_freq <- fcn_freq[order(-fcn_freq$yes_prop),]
  row.names(fcn_freq) <- NULL
  # order plot by frequency
  fcn_df$judge_info <- factor(fcn_df$judge_info, levels = fcn_freq$judge)
  
  # proportion plot
  print(ggplot(fcn_df, aes(y=judge_info, fill = convict)) + 
          geom_bar(position = "fill") + 
          geom_vline(xintercept = mean(fcn_freq$yes_prop), 
                     linetype = "dashed", color = "red") +
          annotate("text", x = -0.02, y = seq_along(fcn_freq$sum), 
                   label = fcn_freq$sum,vjust = 0.5, hjust = +0.7, size = 3) + 
          scale_fill_manual(values = colors_sentences[3:4]) + 
          labs(title = paste(list_state[num_state], "Sentencing Proportions of Top 15 Judges"), x = "Proportion", y = "", fill = "Prison?"))
  
  # count plot
  print(ggplot(fcn_df, aes(y=judge_info, fill = convict)) + geom_bar()+
    annotate("text", x = -0.02, y = seq_along(fcn_freq$yes_prop), label = round(fcn_freq$yes_prop,2),vjust = 0.5, hjust = -0.1, size = 3)+ expand_limits(x = c(0,50))+
    scale_fill_manual(values = colors_sentences[3:4]) +
    labs(title = paste(list_state[num_state], "Sentencing Counts of Top 15 Judges"), x = "Count", y = "", fill = "Prison?")) 
  
  # in time
  print(ggplot(fcn_df_all) + geom_bar(aes(x=as.numeric(as.character(year(Date))), fill = convict), position ="identity", alpha = 0.7) + labs(title = paste("Sentencing Counts Through the Years in",list_state[num_state] ), x = "", subtitle = "All judges", fill = "Prison?")+ scale_x_continuous(n.breaks = 16)+theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)))
}

# want to add average and std dev of proportions, and maybe counts?
#mean(df_freq$D.N.S._prop)
#sd(df_freq$D.N.S._prop)


```

```{r}
num_state = 1
```
# `r list_state[num_state]`

The next 3 plots are repeated for each state. The first plot shows the proportion of prison vs not imprisoned for the top 15 judges for that state, and the number beside each name is the count of cases for that judge. The second plot shows the specific counts, and the number beside each name is the proportion of cases where the accused was sentenced to prison time. The third plot shows the counts per year for all judges.

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r}
judge_counts_prop(num_state)
num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges.

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

# `r list_state[num_state]`

`r list_state[num_state]` has `r nrow(filter(states, state == list_state[num_state]))` cases and `r length(unique(filter(states, state == list_state[num_state])$judge_simple))` unique judges. 

```{r, results='asis'}
judge_counts_prop(num_state)

num_state = num_state + 1 # setting up next cell
```

```{r eval=FALSE}
# EXPORTING AS CSV
write_csv(states, "../data/states_clean.csv")
```

